Thesis Proposal

### 2019-09-18 ThesisProposal.Rmd
Update references and draft presentation

Almost all (social) scientific research copes with missingness (source?). Ignoring missing data can yield wildly invalid inferences (@buur18). A solution is provided by multiple imputation (MI; @rubin87). This technique -- that is growing in popularity (@buur18) -- entails completing a data set with missingness several times. The variability between these completions en-captures the uncertainty of the inferences due to missingness (@rubin87).

The R package MICE (Multiple Imputation using Chained Equations; @mice11) is the world-leading software for multiple imputation (@buur18). However, the package currently does not provide user-friendly means to evaluate the multiply imputed data. Therefore, the goal of this project is develop a MICE evaluation suite, featuring existing modules (like plots to compare the incomplete and completed data sets), and novel assessment tools (like a measure to assess algorithmic convergence). The practical relevance of this research project is in facilitating applied researchers in making valid inferences despite of missingness. Moreover, it contributes to the scientific literature because there is not yet a clear-cut way of diagnosing convergence for MICE algorithms.

@@

During this research project I will work directly with developers of the MICE R package. I will use R Shiny (@shiny17) to program a web-browser based local app for the evaluation of multiply imputed data: 'ShinyMICE'.

@@

A working beta version of ShinyMICE will be considered sufficient to proceed to the next step: writing a technical paper on the workings and usage of the software. Possible journals for publication of this manuscript are *The R Journal* and *Journal of Statistical Software*.

@@

# Additional resources

@@

@abay08

@bart15

@li91

@rubin87

@rubin96

@vinknd

@scha02

@cowl96

@zhu15

@@

Proposal by supervisors:

@@

Bayes course on convergence:
```{r}
# E. Assessing convergence
# -----------------------------------------------------------------------------
# History plot, autocorrelation plot
# History plots show the sampled parameters over the iterations (excluding
# the burn-in).
# The development/pattern in these plots gives an indication of convergence.
# When the history plot is stable (a fat catterpillar), convergence is reached.
# Autocorrelation can be measured at many lags.
# High autocorrelation indicates slow mixing of the random path.

>>>>

At some point, any (social) scientist conducting statistical analyses will run into a missing data problem (@alli02). Simply ignoring the missingness or using ad hoc solutions can yield wildly invalid inferences (@buur18). Multiple imputation (MI; @rubin87) provides a framework to circumvent the problem of missing information. This technique -- that is growing in popularity (@buur18) -- entails completing the missing values in an incomplete data set several times. The variability between the completed data sets captures the uncertainty in the inferences due to missingness (@rubin87).

The R package MICE (Multiple Imputation using Chained Equations; @mice11) is the world-leading software for multiple imputation (@buur18). However, the package currently does not provide user-friendly means to evaluate the multiply imputed data. Therefore, the goal of this project is develop a MICE evaluation suite, featuring existing modules (like plots to compare the incomplete and completed data sets), and novel assessment tools (like a measure to flag algorithmic non-convergence). The practical relevance of this research project is in facilitating applied researchers in making valid inferences despite of missingness. Simultaniously it contributes to the scientific literature because there is not yet a clear-cut way of diagnosing convergence for MI algorithms.

@@

During this research project I will work directly with developers of the MICE R package. I will use R Shiny (@shiny17) to program a web-browser based local app for the evaluation of multiply imputed data: 'ShinyMICE'. This approach does not require the use of any (personal) data. Therefore, the project will not be reviewed by the ethical committee.

@@

A working beta version of ShinyMICE will be considered sufficient to proceed to the next step: writing a technical paper on the workings and usage of the software. The preferred journal for publication of this manuscript is *Journal of Statistical Software* -- alternatively, publication in *The R Journal* will be attempted.

@@

# Additional resources (not used)

@@

<!-- @abay08 -->

<!-- @bart15 -->

<!-- @li91 -->

<!-- @rubin87 -->

<!-- @rubin96 -->

<!-- @vinknd -->

<!-- @scha02 -->

<!-- @cowl96 -->

<!-- @zhu15 -->

@@

**Proposal by supervisors:**

@@

**Bayes course on convergence:**
```{r}
# E. Assessing convergence
# -----------------------------------------------------------------------------
# History plot, autocorrelation plot
# History plots show the sampled parameters over the iterations (excluding
# the burn-in).
# The development/pattern in these plots gives an indication of convergence.
# When the history plot is stable (a fat catterpillar), convergence is reached.
# Autocorrelation can be measured at many lags.
# High autocorrelation indicates slow mixing of the random path.
